---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(pryr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(xxhashlite)
```

# xxhashlite

<!-- badges: start -->
![](https://img.shields.io/badge/cool-useless-green.svg)
[![R-CMD-check](https://github.com/coolbutuseless/xxhashlite/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/coolbutuseless/xxhashlite/actions/workflows/R-CMD-check.yaml)
[![CRAN](https://www.r-pkg.org/badges/version/xxhashlite)](https://CRAN.R-project.org/package=xxhashlite)
<!-- badges: end -->

`xxhashlite` provides access to the *extremely* fast hashing functions
in [xxHash](https://cyan4973.github.io/xxHash/) for in-memory hashing 
of R objects, files, strings and raw vectors.

This package is a wrapper around [xxHash v0.8.2](https://github.com/Cyan4973/xxHash).  
See `inst/LICENSE-xxHash` for the copyright and licensing information for that code.  


## Notes

* Only supports R versions >= v3.5.0 as this is when the serialization 
  infrastructure had a breaking change, and this package will not support
  the old version.

## What's in the box

* `xxhash(robj, algo, as_raw)` calculates the hash of any R object understood by 
`base::serialize()`.
* `xxhash_raw(vec, algo, as_raw)` calculates the hash of a raw vector or string. 
  This function is appropriate when comparing hashes of non-R objects e.g. 
  a checksum hash of raw bytes.
* `xxhash_file(file, algo, as_raw)` calculates the hash of a file
* `xxhash_con(con, algo, as_raw)` calculations the hash of data from a connection
  e.g. `gzfile()`, `url()`

Supports all hashes provided by `xxHash` i.e. XXH128, XXH3, XXH32 and XXH64.


## Installation

This package can be installed from CRAN

``` r
install.packages('xxhashlite')
```

You can install the latest development version from
[GitHub](https://github.com/coolbutuseless/xxhashlite) with:

``` r
# install.package('remotes')
remotes::install_github('coolbutuseless/xxhashlite')
```

Pre-built source/binary versions can also be installed from
[R-universe](https://r-universe.dev)

``` r
install.packages('xxhashlite', repos = c('https://coolbutuseless.r-universe.dev', 'https://cloud.r-project.org'))
```


## Why use a hash?

A hash is a way of succinctly summarising the contents of an object in a compact format.

If there  are changes to the object (no matter how small) then the hash should 
see drastic changes.


```{r}
library(xxhashlite)
xxhash(mtcars)

# Small changes results in a different hash
mtcars$cyl[1] <- 0
xxhash(mtcars)
```

## Available algorithms 
`{xxhashlite}` supports returning the hash as a raw vector, and all the
xxHash algorithms are available:

```{r}
xxhash(mtcars, as_raw = TRUE)
xxhash(mtcars, algo = 'xxh3')
xxhash(mtcars, algo = 'xxh64')
xxhash(mtcars, algo = 'xxh32')
```

## Strings and Raw Vectors

`xxhash()` uses R's serialization mechanism to handle any R object.

When you only require a hash of the raw contents of a file, a raw vector
or a string, then use `xxhash_raw()` and `xxhash_file()`.

Rather than serializing the data first, the hash is calculated on the raw
bytes.




## Acknowledgements

* Yann Collett for releasing, maintaining and advancing [xxHash](https://cyan4973.github.io/xxHash/)
* R Core for developing and maintaining such a great language.
* CRAN maintainers, for patiently shepherding packages onto CRAN and maintaining
  the repository
