---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(dplyr)
library(ggplot2)
library(tidyr)
library(xxhashlite)
```

# xxhashlite

<!-- badges: start -->
![](https://img.shields.io/badge/cool-useless-green.svg)
[![Lifecycle: maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
[![R build status](https://github.com/coolbutuseless/xxhashlite/workflows/R-CMD-check/badge.svg)](https://github.com/coolbutuseless/xxhashlite/actions)
<!-- badges: end -->

`xxhashlite` provides simple access to the *extremely* fast hashing functions
in [xxHash](https://cyan4973.github.io/xxHash/) for in-memory hashing 
of R objects.

This new version of `xxhashlite` (v0.2.0) is a major update which now offers fast
hashing of *any* R object by internally leveraging R's serialization capabilities.

This package is a wrapper around [xxHash v0.8.0](https://github.com/Cyan4973/xxHash).  
See `LICENSE-xxHash`
for the copyright and licensing information for that code.  With this latest
version of xxHash, the new (even faster) hash functions, `xxh3_64bits` and 
`xxhash128`, are considered stable.


## What's in the box

* `xxhash(robj, algo)` calculates the hash of any R object understood by 
`base::serialize()`.


## Installation

You can install from [GitHub](https://github.com/coolbutuseless/xxhashlite) with:

``` r
# install.package('remotes')
remotes::install_github('coolbutuseless/xxhashlite)
```

#### Installation - set CFLAGs for optimised executable

To get the most out of what `xxHash` offers, it will be important to set
some optimization flags for your machine.  The important compiler flags to set 
are `-O3` and `-march=native`.

Here are 2 possible ways to do this:

1. Copy `src/Makevars.custom` to `src/Makevars` re-build package.
2. Edit your `~/.R/Makevars` to include `CFLAGS = -O3 -march=native` (this will
  change flags for all future compilation, and should probably be used with caution)


## Why use a hash?

A hash is a way of succinctly summarising the contents of an object in a compact format.

If there  are changes to the object (no matter how small) then the hash should change as 
well.


```{r}
library(xxhashlite)
xxhash(mtcars)

# Small changes results in a different hash
mtcars$cyl[1] <- 0
xxhash(mtcars)
```



## Timing for hashing arbitrary R objects

`xxhashlite` uses the [xxHash](https://github.com/Cyan4973/xxHash) family 
of hash functions to provide very fast hashing of R objects.

For the test case shown below, `xxhashlite` is faster at calculating a hash 
than all other methods I could find, with a maximum hashing speed in this
specific case of 20 GB/s.

Note: actual hashing speed will still depend on R's serialization functions 
e.g. small complex data.frames might have a lot of serialization overhead compared
to long numeric vectors.


<details>
<summary> Click to show/hide the benchmarking code </summary>

```{r bench1}
library(xxhashlite)
library(digest)
library(fastdigest)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Simple data.frame
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
N   <- 1e5
df  <- data.frame(
  x = runif(N),
  y = sample(N)
)

size <- pryr::object_size(df)
size


res <- bench::mark(
  # {xxhashlite}
  xxhash(df, 'xxhash32'),
  xxhash(df, 'xxhash64'),
  xxhash(df, 'xxhash128'),
  xxhash(df, 'xxh3_64bits'),
  
  # {digest}
  digest(df, algo = 'xxhash32'),
  digest(df, algo = 'xxhash64'),
  digest(df, algo = 'murmur32'),
  digest(df, algo = 'spookyhash'),
  
  # {fastdigest}
  fastdigest(df),
  
  check = FALSE
)
```

</details>


```{r echo = FALSE}
res %>% 
  mutate(`MB/s` = round(size/1024^2 / as.numeric(median), 1)) %>%
  mutate(`itr/sec` = round(`itr/sec`)) %>%
  select(expression, median, `itr/sec`, `MB/s`) %>%
  mutate(package = c(
    # rep('xxhashlite - data only',  4),
    rep('xxhashlite'            ,  4),
    rep('digest'                ,  4),
    rep('fastdigest'            ,  1)
  )) %>%
  select(package, everything()) %>%
  knitr::kable(caption = "Hashing a simple data.frame")
```


```{r echo = FALSE}
plot(res) + 
  theme_bw()
```


## Related Software

* The original [xxHash](https://cyan4973.github.io/xxHash/) software.
* The [`digest`](https://cran.r-project.org/package=digest) package is 
  a much more comprehensive approach to hashing from within R - multiple 
  hashing functions and much more configurability
* The [`hashFunction`](https://cran.r-project.org/package=hashFunction) package offers 
  some hasing functions that work directly on character or integer values.
* The [`fastdigest`](https://cran.r-project.org/package=fastdigest) offers
  the fast non-cryptographic 'SpookyHash' and will hash anything that
  serialize knows about.

## Acknowledgements

* Yann Collett for releasing, maintaining and advancing [xxHash](https://cyan4973.github.io/xxHash/)
* R Core for developing and maintaining such a great language.
* CRAN maintainers, for patiently shepherding packages onto CRAN and maintaining
  the repository
