---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(dplyr)
library(xxhashlite)
```

# xxhashlite

<!-- badges: start -->
![](https://img.shields.io/badge/cool-useless-green.svg)
<!-- badges: end -->

`xxhashlite` provides simple access to the *extremely* fast hashing functions
in [xxHash](https://cyan4973.github.io/xxHash/) for in-memory hashing 
of R atomic, numeric vectors.

The scope of this package is limited - it aims to provide functions for direct,
in-memory hashing of vectors which contain raw, integer, real, complex or logical values.  

If you wanted to hash arbitrary R objects, you could first turn it 
into a raw vector representation
using `base::serialize()`. 

Currently xxHash code provided with this package is v0.7.3.  See `LICENSE-xxHash`
for the copyright and licensing information for that code.


### Design choices

* `xxhashlite` will hash the *data payload* within an object, and not the R 
  object itself.  This means that hashes calculated within R will match 
  hashes calculated on the equivalent data in a file hashed by the command 
  line programs provided with `xxHash`.

### Limitations

* As it is the *data payload* of the vector or 
  matrix that is being hashed, this
  does not include any notion of the container for that data.  This
  means that a vector and array which contain the same data will hash to the 
  same value - regardless of the dimensions or other attributes.
* `xxHash v0.7.x` includes the experimental `xxh3` and `xxh128` hash functions.
   From the documentation: "The algorithm is currently in development, meaning 
   its return values might still change in future versions. However, the API is 
   stable, and can be used in production, typically for generation of 
   ephemeral hashes (produced and consumed in same session)".
* `xxHash` is a non-cryptographic hash.


## Installation

You can install from [GitHub](https://github.com/coolbutuseless/xxhashlite) with:

``` r
# install.package('remotes')
remotes::install_github('coolbutuseless/xxhashlite)
```

#### Installation - set CFLAGs for optimised executable

To get the most out of what `xxHash` offers, it will be important to set
some optimization flags for your machine.  The important compiler flags to set 
are `-O3` and `-march=native`.

Here are 2 possible ways to do this:

1. Copy `src/Makevars.custom` to `src/Makevars` re-build package.
2. Edit your `~/.R/Makevars` to include `CFLAGS = -O3 -march=native` (this will
  change flags for all future compilation, and should probably be used with caution)




## Simple hashing of raw, integer, real, complex and logical vectors

The supplied functions operate directly on the payload within
vectors, matrices or arrays containing raw, integer, real, complex or logical values.

Because these functions operate on the raw data contained in the vector and not
the R object itself, the attributes and dimensions of the object
are not included in the hash.

By not having to serialize the R object first, overall hashing time is 
reduced.  The price paid for this however is that a vector and matrix with the 
same underlying values, but different dimensions, will hash to the same value.

Four hash functions from xxHash's "simple api" are exposed:

* **xxhash32** - 32 bit output. Will be slow on a 64bit machine.
* **xxhash64** - 64 bit output. About 2x faster than xxhash32
* **xxhash128** - 128 bit output. **Marked as experimental in xxHash**
* **xxh3_64bits** - 64 bit output. **Marked as experimental in xxHash**




```{r}
library(xxhashlite)

vec <- raw(1e6)
mat <- as.array(vec, 1e3, 1e3)

xxhashlite::xxhash32(vec)
xxhashlite::xxhash32(mat)

xxhashlite::xxhash64(vec)
xxhashlite::xxhash64(mat)

xxhashlite::xxhash128(vec)
xxhashlite::xxhash128(mat)

xxhashlite::xxh3_64bits(vec)
xxhashlite::xxh3_64bits(mat)
```


## Hashing 1 million raw bytes

An R vector of raw bytes may be hashed as:

* An R object containing opaque data e.g. Using `xxhash32()`
* The raw data values contained in the object e.g. Using `xxhash32()`

Hashing the raw data in the object is 5x to 20x faster than hashing the entire object.
The reason for this speed difference is because when hashing the entire R object, 
it must first be converted to a raw byte representation (using `base::serialize()`) - 
and this conversion operation is expensive.

When hashed as just the data payload, `xxHash` has a throughput of around ~25 GB/s 
(2015 MacBook Pro), but your mileage will vary depending on your CPU and 
compiler flags.


<details>
<summary> Click here to show/hide benchmark code </summary>

```{r bench1}
library(xxhashlite)

N   <- 1e6
vec <- as.raw(seq(N) %% 256)

res <- bench::mark(
  xxhash32(vec),
  xxhash64(vec),
  xxhash128(vec),
  xxh3_64bits(vec),
  check = FALSE
)
```

</details>


```{r echo = FALSE}
res %>% 
  mutate(`GB/s` = round(N/1024^3 / as.numeric(median), 1)) %>%
  mutate(`itr/sec` = round(`itr/sec`)) %>%
  select(expression, median, `itr/sec`, `GB/s`) %>%
  mutate(package = c(
    rep('xxhashlite', 4)
  )) %>%
  select(package, everything()) %>%
  knitr::kable(caption = "Hashing 1 million raw bytes")
```


## Hashing arbitrary R objects

To hash arbitrary R objects (including numeric vectors with dimensions and other attributes),
use `base::serialize()` to convert the object to a stream of raw bytes.


```{r}
xxhash64(base::serialize(mtcars, NULL))
```


## Related Software

* The original [xxHash](https://cyan4973.github.io/xxHash/) software.
* The [`digest`](https://cran.r-project.org/package=digest) package is 
  a much more comprehensive approach to hashing from within R - multiple 
  hashing functions and much more configurability
* The [`hashFunction`](https://cran.r-project.org/package=hashFunction) package offers 
  some hasing functions that work directly on character or integer values.
* The [`fastdigest`](https://cran.r-project.org/package=fastdigest) offers
  the fast non-cryptographic 'SpookyHash' and will hash anything that
  serialize knows about.

## Acknowledgements

* Yann Collett for releasing, maintaining and advancing [xxHash](https://cyan4973.github.io/xxHash/)
* R Core for developing and maintaining such a great language.
* CRAN maintainers, for patiently shepherding packages onto CRAN and maintaining
  the repository
